Computing in the Field of Rationals

### 2012-06-21 rational_ext.tex
_

\begin{definition}[Size of a type]
\label{def:size}
The size of a type {{b}} is a number defined to be:
\vspace{-20pt}
\begin{multicols}{2}
%subcode{opsem} include main
% [^ b1 + b2 ^] '= [^ b1 ^] + [^ b2 ^]
% [^ b1 * b2 ^] '= [^ b1 ^] * [^ b2 ^]

@@

\begin{lemma}[{{Pi}} is a category]
  The category {{Pi}} has the types {{b}} as objects and equivalence
  classes of well-typed combinators {{b1<->b2}} as morphisms. One can
  check:
  \begin{enumerate}
  \item Every object {{b}} has an identity morphism {{id : b <->b}}.
  \item Composition {{g (o) f}} of morphisms {{f:b1 <->b2}} and
    {{g:b2<->b3}} is given by sequencing {{f (;) g}}. 
  \item Associativity of composition follows from operational
    equivalence of {{f(;)(g(;)h)}} and {{(f(;)g)(;)h}} (where
    {{h:b3<->b4}}). One can check:

@@

    \begin{enumerate}
    \item The braiding must satisfy two ``hexagon axioms''.
    \item The braiding is self inverse,
      {{swap+_{b1+b2}=(swap+_{b2+b1})^{dagger} }}.
    \end{enumerate}

@@

\noindent
No division by zero.

@@

%%%%%%%%%%%%%%%%%%
\subsection{Cloning}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Construction of the Solver}

>>>>

\begin{definition}[Size of a type]
\label{def:size}
The size of a type {{b}}, denoted by {{[^ b ^]}}, is a numeric value
and is defined to be:
\vspace{-20pt}
\begin{multicols}{2}
%subcode{opsem} include main
% [^ b1 + b2 ^] '= [^ b1 ^] + [^ b2 ^]
% [^ b1 * b2 ^] '= [^ b1 ^] * [^ b2 ^]

@@

\begin{lemma}[{{Pi}} is a category]
  The category {{Pi}} has the types {{b}} as objects and equivalence
  classes of well-typed combinators {{b1<->b2}} as morphisms. One can
  check:
  \begin{enumerate}
  \item Every object {{b}} has an identity morphism {{id : b <->b}}.
  \item Composition {{g (o) f}} of morphisms {{f:b1 <->b2}} and
    {{g:b2<->b3}} is given by sequencing {{f (;) g}}. 
  \item Associativity of composition follows from operational
    equivalence of {{f(;)(g(;)h)}} and {{(f(;)g)(;)h}} (where
    {{h:b3<->b4}}).

    Assuming {{v1:b1}}, {{v2:b2}}, {{v3:b3}}, {{v4:b4}},
    {{f~v1|-->v2}}, {{g v2 |--> v3}} and {{h~v3 |--> v4}}, one can
    check:

@@

    \begin{enumerate}
    \item The braiding must satisfy two ``hexagon'' axioms.
    \item The braiding is self inverse,
      {{swap+_{b1+b2}=(swap+_{b2+b1})^{dagger} }}.
    \end{enumerate}

@@

\paragraph*{No division by zero.} 
The important side-condition on the types not captured by the syntax
above is that there are no types of the form {{1/b}}, where
{{[^b^]=0}} (see Defn. \ref{def:size}). This eliminates syntaxes of
the form {{1/0}}, {{1/(0+0)}}, {{1/(0*b)}} etc, from being treated as
valid {{b}} types. Intuitively, this translates into the familiar
restriction that there is no division by 0 in arithmetic or the
definition of fields in abstract algebra.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cloning}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Construction of the Solver}
